---
title: "Most Streamed Spotify Songs 2024"
authors: "Peter Quiros and Jennifer Cajas"
format: html
editor: visual
---

# Data

[Data dictionary](https://www.kaggle.com/datasets/nelgiriyewithana/most-streamed-spotify-songs-2024)

-   **Track Name**: Name of the song.

-   **Album Name**: Name of the album the song belongs to.

-   **Artist**: Name of the artist(s) of the song.

-   **Release Date**: Date when the song was released.

-   **ISRC**: International Standard Recording Code for the song.

-   **All Time Rank**: Ranking of the song based on its all-time popularity.

-   **Track Score**: Score assigned to the track based on various factors.

-   **Spotify Streams**: Total number of streams on Spotify.

-   **Spotify Playlist Count**: Number of Spotify playlists the song is included in.

-   **Spotify Playlist Reach**: Reach of the song across Spotify playlists.

-   **Spotify Popularity**: Popularity score of the song on Spotify.

-   **YouTube Views**: Total views of the song's official video on YouTube.

-   **YouTube Likes**: Total likes on the song's official video on YouTube.

-   **TikTok Posts**: Number of TikTok posts featuring the song.

-   **TikTok Likes**: Total likes on TikTok posts featuring the song.

-   **TikTok Views**: Total views on TikTok posts featuring the song.

-   **YouTube Playlist Reach**: Reach of the song across YouTube playlists.

-   **Apple Music Playlist Count**: Number of Apple Music playlists the song is included in.

-   **AirPlay Spins**: Number of times the song has been played on radio stations.

-   **SiriusXM Spins**: Number of times the song has been played on SiriusXM.

-   **Deezer Playlist Count**: Number of Deezer playlists the song is included in.

-   **Deezer Playlist Reach**: Reach of the song across Deezer playlists.

-   **Amazon Playlist Count**: Number of Amazon Music playlists the song is included in.

-   **Pandora Streams**: Total number of streams on Pandora.

-   **Pandora Track Stations**: Number of Pandora stations featuring the song.

-   **Soundcloud Streams**: Total number of streams on Soundcloud.

-   **Shazam Counts**: Total number of times the song has been Shazamed.

-   **TIDAL Popularity**: Popularity score of the song on TIDAL.

-   **Explicit Track**: Indicates whether the song contains explicit content.

```{r}
library(tidyverse)
songs <- read_csv("data/Most Streamed Spotify Songs 2024.csv")
```

# Questions

Do tracks with explicit lyrics have higher streams than ones that don't?

Is there a relationship between Youtube Views and Track Score?

# Descriptive Analysis

## Histogram

What is the distribution of the number of streams that are explicit?

```{r}
songs |>
  group_by(`Explicit Track`) |>
  ggplot(aes(x = `Spotify Streams`, na.rm = TRUE)) +
  geom_histogram(bins = 50, na.rm = TRUE)
```

## Box plot

What is the distribution of the number of streams that are explicit?

```{r}
songs |>
  ggplot(aes(x = `Explicit Track`, y = `Spotify Streams`)) +
  geom_boxplot()
```

## Scatter Plot

Is there a relationship between Youtube Views and Spotify Streams?

```{r}
songs |>
  ggplot(aes(x = `YouTube Views`, y = `Spotify Streams` )) +
  geom_point()
```

## Data Summariziation

```{r}
songs |>
  summarize(mean = mean(`Spotify Streams`, na.rm = TRUE),
            sd = sd(`Spotify Streams`, na.rm = TRUE),
            min = min(`Spotify Streams`, na.rm = TRUE),
            max = max(`Spotify Streams`, na.rm = TRUE))

```

# Inferential Statistics

## T-test

Do tracks with explicit lyrics have higher streams than ones that don't?

```{r}
t.test(`Spotify Streams` ~ `Explicit Track`, 
       data = songs)
```

## Linear Regression

Is there a relationship between Youtube Views and Track Score?

```{r}
yt_score <- lm(`YouTube Views` ~ `Track Score`,
               data = songs)
summary(yt_score)
```

# Conclusion

Using a Histogram to answer the following question of What is the distribution of the number of streams that are explicit? The histogram showed that the number of explicit streams shows a right-skewed pattern, indicating that the most values are concentrated on the lower end, with fewer streams having very high counts. Using the t-Test the two values shown  of 0.04127 and p-value of 0.9671 indicate that there is not much of a difference in the average number of streams between explicit and non-explicit songs. Since the p-value is much greater than 0.05, we fail to reject the null hypothesis and conclude that any observed difference is likely due to chance.

The linear regression shows that Track Score is a significant factor in predicting YouTube Views (p \< 2e-16). For every 1-point increase in Track Score, views go up by about 3.86 million on average. But the R-squared value is only 0.044, meaning Track Score explains just 4.4% of the changes in views. While the relationship is significant, it is not strong suggesting that other factors likely have a larger impact on view counts. 
